---
title: "Project 2"
authors: "Noel Pop and Georgia Cameron"
date: "10/31/2025"
output:
  pdf_document
subtitle: "CPSC 292 Section #02#"
---

# Project 1: Best-Selling Video Games of All Time

```{r setup, include = FALSE}
#Loading required libraries
library(tinytex)
library(tidyr)
library(janitor)
library(readxl)
library(lubridate)
library(ggplot2)
library(dplyr)
library(paletteer)
library(patchwork)
```

## Loading And Cleaning Data
```{r}
game.data <- read.csv("game.data/best-selling video games of all time.csv")
game.data
```

```{r}
game.data.clean <- game.data %>%
clean_names() %>%
remove_empty(which = c("rows","cols")) %>%
na.omit(game.data) %>%
mutate(sales = sales / 100000)
game.data.clean
```
### Cleaning Steps Description:
This cleans the data names, removes any empty columns or rows, and omits any NA data.

## Visualizations

### Original Plot 1:
```{r}
plot1 <- ggplot(game.data.clean) + geom_boxplot(aes(x = platform_s, y = sales, fill = platform_s)) + theme(legend.position = "none", axis.title.x = element_text(face = "bold"), axis.title.y = element_text(face = "bold"), plot.title = element_text(face = "bold")) + labs(title = "Number Of Sales of Best-Selling Games On Each Platform", x = "Platform", y = "Number of Sold Copies (x100000)") + scale_fill_paletteer_d("MetBrewer::Signac") + coord_flip()
plot1
```
### Description of Plot 1:
This plots shows us different platforms that the most sold video games of all time uses.

### Original Plot 2:
```{r}
color.base <- paletteer::paletteer_d("MetBrewer::Signac")
color <- rep(color.base, length.out = 18)
plot2 <- ggplot(game.data.clean) + geom_bar(aes(x = publisher_s, fill = publisher_s)) + theme(legend.position = "none", axis.title.x = element_text(face = "bold"), axis.title.y = element_text(face = "bold"), plot.title = element_text(face = "bold")) + labs(title = "How Many Best-Selling Games Sold By Each Publisher", x = "Publisher", y = "Number of Games", fill = "Publisher") + scale_fill_manual(values = color, name = "Publisher") + coord_flip()
plot2
```
### Description of Plot 2:
These plots show us how many of the best selling video games of all time belong to each publisher.


### Plot 1 with Function

```{r}
make_plot <- function(data, x_var, y_var, geom_layer) {
  ggplot(data, aes_string(x = x_var, y = y_var, fill = x_var)) + 
    geom_layer() +
    theme(
      legend.position = "none",
      axis.title.x = element_text(face = "bold"),
      axis.title.y = element_text(face = "bold"),
      plot.title = element_text(face = "bold")
    ) +
    coord_flip()
}

plot1 <- make_plot(game.data.clean, "platform_s", "sales", geom_boxplot) 

plot1 <- plot1 + 
  labs(
      title = "Number Of Sales of Best-Selling Games On Each Platform", 
      x = "Platform", 
      y = "Number of Sold Copies (x100000)") +
  scale_fill_paletteer_d("MetBrewer::Signac")


plot1
```

## Description of Plot 1 with Function
This plot produces the same visualization as the original code. A "make_plot" function was made to simplify the code and accept arguments including the data set, x variable, y variable, and geom layer style. 

### Plot 2 with Function and Conditional

```{r}

plot2 <- make_plot(game.data.clean, "publisher_s", geom_layer = geom_bar) +
  labs(
    title = "How Many Best-Selling Games Sold By Each Publisher",
    x = "Publisher",
    y = "Number of Games",
    fill = "Publisher"
  ) + 
  scale_fill_manual(values = color) +
  geom_text(
    stat = "count",
    aes(label = ifelse(after_stat(count) > 1, after_stat(count), "")),
    vjust = 0,
    hjust = -0.5,
    size = 2
  )

plot2
```
## Description of Plot 1 with Function
This plot produces the same visualization as the original code. The "make_plot" function was used to create the visualization. A conditional was added to the geom text layer to include labels for the number of games sold by each publisher if the value was greater than 1. This feature adds an additional visual element that makes interpreting the data easier. 


## References
https://dplyr.tidyverse.org/
https://ianadamsresearch.com/post/using-ggplot2-to-visualize-the-frequency-of-your-name/
https://github.com/thomasp85/patchwork#patchwork
https://www.sthda.com/english/wiki/bar-plots-r-base-graphs
https://exts.ggplot2.tidyverse.org/gallery/ 
https://r-graph-gallery.com/ggplot2-color.html




